\plaintitle

### 2016-05-29 paper/paper.tex
Fixed citations mostly
Pretty decent now

\documentclass[chi_draft]{sigchi}
% Use this command to override the default ACM copyright statement
% (e.g. for preprints).  Consult the conference website for the
% camera-ready copyright statement.

@@

\usepackage[backend=biber, style=numeric, firstinits=true, url=false, sorting=none]{biblatex}
%\usepackage{custom}

@@

\begin{abstract}
\todo{Abstracts should be about 150 words}
    Visualization of sports data is very common in recent years as access to
    data has increased. The abundance of data and audience allows for new and
    innovative presentations. In this paper we describe an interactive
    visualization of the NBA competition over the last 30 years. With three
    different views, we allow the user to explore data about NBA seasons
    including standings, team statistics, transfers and player statistics. The
    d3.js framework was used to develop this visualization. 
\end{abstract}

@@

The data was gathered by scraping the basketball-reference site using the
provided download capabilities. Most of the data was downloaded in csv format,
while some tables had to be manually scraped. This process was automated using
Python scripts. The data was then combined in a preprocessing step. In this
step, each team's playoff rankings were calculated based on the matches played
during the playoffs, and the rest of the data was combined into json format. The
final preprocessing step combined all data into one json file. A model of the
available data can be seen on figure~\ref{fig:data} \todo{is die afbeelding nog wel juist? Is die nuttig?}.

@@

Our initial play-off view was too dense because of the connections between teams
as well. Teams that played against each other were connected directly. When
searching for alternatives to reflect on our choice we discovered a similar
visualization in a totally different context\cite{whitehousepath} as explained in 
the section \ref{sec:literature}. With the idea of this visualization in mind, our 
visualization gives a cleaner view of the competitors for a specific team compared 
to our original sketch. Hence we decided to adapt our visualization. We added an 
intermediate step between two teams to more clearly indicate how teams competed to 
become the NBA champion. \todo{meaning of last sentence is not clear}

\todo{Hierarchy = sorting = facilitate trends of winners
position have a meaning in the play-off view
when selecting circle => zoom/more detail
Bubble principle of 
- Symmetry (whole = play-off)
- conectedness arc = games played
Pre attentive characteristics
- size of circles
- hue to highligt selections or indicate increase decline arrows next view}

@@

\subsubsection{Arrow section}
This section illustrates a change for a chosen statistic for the selected team in 
the selected season compared to the previous season. This gives a user more 
insight into why a statistic changed over time and how this change has affected 
team performance. The selected team is represented by a circle in the same way as 
explained in the play-off view. To give the user a better interpretation of the SRS 
value of the circle, three other reference circles are drawn. These are 
transparant with a lightgrey border. They show the maximum, average and minimum SRS 
value of the NBA in the selected season. The change for the chosen statistic is 
represented by three arrows:
\begin{itemize}
    \item Arrow on the left pointing towards the circle: indicating the
        influence of players who joined the team.
    \item Arrow on the right of the circle: indicating how the team internally
        changed, e.g. by training its players
    \item Arrow on the left pointing away from the circle: indicating the
        influence of players who left the team 
\end{itemize}
\todo{add statistics view image here}
The value of the change in statistic is shown next to the arrow. The number of 
players responsible for the change is shown between brackets. Their are shown 
jerseys next to the arrows as well. These are the jerseys of the two players that 
had the biggest contribution to the influence represented by that arrow.  
On the left, there is a possibility to chose a statistic. The possibilities 
are the PER value, the age, the effective field goal percentage, the offensive rating 
and the defensive rating. Below this, player information is shown when a user hovers
over a player's jersey.

@@

\subsubsection{Previous designs}
Initially we designed this view more elaborate. When the user would click on a
jersey, specific player information would be shown together with a number of player
statistics in the form of small multiples. A bell curve should have informed
how a player scores compared to his team, other players in the league on his
position or all other players in the league. We considered other charts for the
small multiples as well. A box plot or bullet was less clear for us to indicate
where a player stands compared to the group he was compared with. This part of the 
visualization was not our main focus. Because of the lesser importnace and because of time and resource constraints, it has not been fully developed. \todo{sketch of small multiples
here}

\subsection{Improving the result}\todo{Is this necessary? I don't know what to
write here.}
Talk about how we improved the data-ink ration in the end..

\subsection{Technology}
To create the visualization, the d3 javascript framework~\cite{d3} was used in
combination with html5~\cite{html5} and jquery~\cite{jquery}. Note that D3 is
based on HTML, CSS, Javascript and SVG. This allows easy access to the
visualization as most modern browsers are capable of handling these
webstandards. The choice not to use the d3 framework for the entire
visualization was made to ease the layout configuration of the visualization.
Instead, html was used to do the global layout of the visualization. We opted to
structure the site with multiple divs as containers. To each container we
allocate a svg with a specific visualization.
\todo{sketch of how the page is structured here}

@@

\todo{References should be in alphabetical order and represented by numbers in the text. Style should be set to style{SIGCHI-Reference-Format}, this moment it is set to the style draft}

>>>>

\documentclass[]{sigchi}
% Use this command to override the default ACM copyright statement
% (e.g. for preprints).  Consult the conference website for the
% camera-ready copyright statement.

@@

\usepackage[backend=biber, style=numeric, sorting=nty, firstinits=true, url=false]{biblatex}
%\usepackage{custom}

@@

\begin{abstract}
    Visualization of sports data is very common in recent years as access to
    data has increased. The abundance of data and audience allows for new and
    innovative presentations. In this paper we describe an interactive
    visualization of the NBA competition over the last 30 years. With three
    different views, we allow the user to explore data about NBA seasons
    including standings, team statistics, transfers and player statistics. The
    d3.js framework was used to develop this visualization. 
\end{abstract}

@@

The data was gathered by scraping the basketball-reference site using the
provided download capabilities. Most of the data was downloaded in csv format,
while some tables had to be manually scraped. This process was automated using
Python scripts. The data was then combined in a preprocessing step. In this
step, each team's playoff rankings were calculated based on the matches played
during the playoffs, and the rest of the data was combined into json format. The
final preprocessing step combined all data into one json file. A model of the
available data can be seen on figure~\ref{fig:data}.

@@

Our initial play-off view was too dense because of the connections between teams
as well. Teams that played against each other were connected directly. When
searching for alternatives to reflect on our choice we discovered a similar
visualization in a totally different context\cite{whitehousepath} as explained in 
the section \ref{sec:literature}. With the idea of this visualization in mind, our 
visualization gives a cleaner view of the competitors for a specific team compared 
to our original sketch. Hence we decided to adapt our visualization. We added an 
intermediate connection point between two teams to more clearly indicate how teams competed to 
become the NBA champion.

%\todo{Hierarchy = sorting = facilitate trends of winners
%position have a meaning in the play-off view
%when selecting circle => zoom/more detail
%Bubble principle of 
%- Symmetry (whole = play-off)
%- conectedness arc = games played
%Pre attentive characteristics
%- size of circles
%- hue to highligt selections or indicate increase decline arrows next view}

@@

\subsubsection{Arrow section}
This section illustrates a change for a chosen statistic for the selected team in 
the selected season compared to the previous season. This gives a user more 
insight into why a statistic changed over time and how this change has affected 
team performance. The selected team is represented by a circle in the same way as 
explained in the play-off view. To give the user a better interpretation of the SRS 
value of the circle, three other reference circles are drawn. These are 
transparant with a lightgrey border. They show the maximum, average and minimum SRS 
value of the NBA in the selected season. The change for the chosen statistic is 
represented by three arrows:
\begin{itemize}
    \item Arrow on the left pointing towards the circle: indicating the
        influence of players who joined the team.
    \item Arrow on the right of the circle: indicating how the team internally
        changed, e.g. by training its players
    \item Arrow on the left pointing away from the circle: indicating the
        influence of players who left the team 
\end{itemize}
The value of the change in statistic is shown next to the arrow. The number of 
players responsible for the change is shown between brackets. Their are shown 
jerseys next to the arrows as well. These are the jerseys of the two players that 
had the biggest contribution to the influence represented by that arrow.  
On the left, there is a possibility to chose a statistic. The possibilities 
are the PER value, the age, the effective field goal percentage, the offensive rating 
and the defensive rating. Below this, player information is shown when a user hovers
over a player's jersey.

@@

\subsubsection{Previous designs}
Initially we designed this view more elaborate. When the user would click on a
jersey, specific player information would be shown together with a number of
player statistics in the form of small multiples. A bell curve should have
informed how a player scores compared to his team, other players in the league
on his position or all other players in the league. We considered other charts
for the small multiples as well. A box plot or bullet was less clear for us to
indicate where a player stands compared to the group he was compared with. This
part of the visualization was not our main focus. Because of the lesser
importance and because of time and resource constraints, it has not been fully
developed. A sketch including the small multiples view on the bottom right is
shown on Figure~\ref{fig:smallmultsketch}.

\begin{figure}
\centering
  \includegraphics[width=1.0\columnwidth]{figures/smallmultsketch}
  \caption{A sketch of a previous design including the small multiples on the
  bottom right.}
  \label{fig:smallmultsktech}
\end{figure}

\subsection{Technology}
To create the visualization, the d3 javascript framework~\cite{d3} was used in
combination with html5~\cite{html5} and jquery~\cite{jquery}. Note that D3 is
based on HTML, CSS, Javascript and SVG. This allows easy access to the
visualization as most modern browsers are capable of handling these
webstandards. The choice not to use the d3 framework for the entire
visualization was made to ease the layout configuration of the visualization.
Instead, html was used to do the global layout of the visualization. We opted to
structure the site with multiple divs as containers. To each container we
allocate a svg with a specific visualization.
%\todo{sketch of how the page is structured here}

@@
